---
title: "Integrated step selection analysis (iSSA)"
author: "Pierre Cottais & An Hoàng"
date: "25/01/2022"
output: 
  pdf_document: 
    extra_dependencies: ["float"]
    toc: yes
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide")
```


```{r packages}
library(tidyverse)
library(ggmap)
# library(sf)
library(lubridate)
library(amt)
library(stargazer)
```


# Data processing
```{r}
# oceanography dataset
load("CALDIO_MPE_2021/generalisation datasets/grid_oceano_2011.Rdata")
grid_oceano_2011 <- grid_oceano
load("CALDIO_MPE_2021/generalisation datasets/grid_oceano_2012.Rdata")
grid_oceano_2012 <- grid_oceano
remove(grid_oceano)

# related dates
load("CALDIO_MPE_2021/generalisation datasets/grid_oceano_Date_2011.Rdata")
grid_Oceano_Date_2011 <- grid_Oceano_Date
load("CALDIO_MPE_2021/generalisation datasets/grid_oceano_Date_2012.Rdata")
grid_Oceano_Date_2012 <- grid_Oceano_Date
remove(grid_Oceano_Date)

# Scopoli's sheawater datasets
load("CALDIO_MPE_2021/training datasets/CALDIO_2011_oceano.Rdata")
load("CALDIO_MPE_2021/training datasets/CALDIO_2012_oceano.Rdata")
ALL <- bind_rows(ALL2011, ALL2012)
ALL <- ALL %>% mutate(Year = as.factor(year(ymd_hms(Time))))
# remove(ALL2011, ALL2012)


# load functions
source("functions.R")
```


Work with Shoreline / Coastline shapefiles
```{r eval=FALSE, include=FALSE}
# europe <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf",
#                                      continent = "Europe")
# 
# mediterra <- europe %>% 
#   filter(admin %in% c("France", "Italy", "Spain")) %>%  
#   st_crop(xmin = 3, xmax = 11, ymin = 40, ymax = 45)
# 
# mediterra %>% 
#   # st_transform(crs = 2154) %>% 
#   ggplot() + geom_sf() + 
#   geom_point(data = grd[!is.na(grd$Bathy) & !is.na(grd$SST1),],
#              aes(x = Longitude, y = Latitude), size = 0.2)
# 
# df_union_cast <- mediterra$geometry %>% 
#   st_cast("MULTIPOLYGON")
# 
# plot(df_union_cast)
```

Remove cells near the coast boundary
```{r message=FALSE, warning=FALSE}
grd <- grid_oceano_2011[[1]]
mediterranean  <- make_bbox(lat = Latitude, lon = Longitude, data = grd)
map <- get_map(location = mediterranean) %>% ggmap()
df_grid <- grd %>% filter(!is.na(SST1), !is.na(Bathy))
map + geom_point(data = df_grid, aes(x = Longitude, y = Latitude), size = 0.1)
```


Visualise colonies
```{r eval=FALSE, include=FALSE}
colony_colors <- c("#FC8D58", "#D72F28", "#1A7838", "#69ADDB")
png("figures/map_sites.png", width=7, height=4.2, units="in", res=200)
mediterranean  <- make_bbox(lat = Latitude, lon = Longitude, data = ALL2011)
map <- get_stamenmap(bbox = mediterranean, maptype = "toner-lite", zoom = 7) %>% ggmap()
map + geom_point(data = ALL2011, aes(x = Longitude, y = Latitude, color = Site), size = 0.1) +
  scale_color_manual(values = colony_colors) +
  xlab("Longitude") + ylab("Latitude") +
  labs(title = "Répartition des individus selon le site de leur colonie en 2011")+
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  guides(colour = guide_legend(override.aes = list(size=3)))
dev.off()
```

```{r eval=FALSE, include=FALSE}
# colony_colors <- c("#FC8D58", "#D72F28", "#1A7838", "#69ADDB")
# png("figures/bar_site.png", width=8, height=3, units="in", res=200)
# ALL2011 %>% group_by(Site) %>% 
#   summarise(nb_ind = n_distinct(ID)) %>% 
#   ggplot() + geom_col(aes(x = Site, y = nb_ind), width = 0.5,
#                       fill = colony_colors) +
#   ylab(element_blank()) + xlab(element_blank()) +
#   labs(title = "Répartition des individus selon le site de leur colonie en 2011",
#        caption = "Nombre total d'individus : 94") +
#   theme(plot.title = element_text(size = 16, hjust = 0.5, vjust = 2.7),
#         plot.caption = element_text(size = 10, face = "italic", hjust = 1),
#         legend.position = "bottom",
#         legend.title = element_blank(),
#         axis.text.x = element_text(vjust = 2, size = 14, color = colony_colors),
#         axis.text.y = element_text(size = 10),
#         panel.background = element_rect(fill = "white"),
#         panel.grid = element_line(color = 'grey', linetype = 'dotted'),
#         panel.grid.major.x = element_blank(),
#         axis.ticks.x = element_blank())
# dev.off()
```


```{r eval=FALSE, include=FALSE}
colony_colors <- c("#FC8D58", "#D72F28", "#1A7838", "#69ADDB")
png("figures/bar_site.png", width=8, height=3, units="in", res=200)
ALL %>% 
  # mutate(ID = paste(ID, Year, sep = '')) %>% 
  group_by(Year, Site) %>% 
  summarise(nb_ind = n_distinct(ID)) %>% 
  ggplot() + aes(x = Site, y = nb_ind, fill = Year) + 
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8, width = 0.5) +
  ylab(element_blank()) + xlab(element_blank()) +
  labs(title = "Répartition des individus selon le site de leur colonie",
       caption = "Nombre total d'individus par année : 94") +
  theme(plot.title = element_text(size = 16, hjust = 0.5, vjust = 2.7),
        plot.caption = element_text(size = 10, face = "italic", hjust = 1),
        legend.position = "right",
        legend.title = element_blank(),
        axis.text.x = element_text(vjust = 2, size = 14, color = colony_colors),
        axis.text.y = element_text(size = 10),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(color = 'grey', linetype = 'dotted'),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_blank())
dev.off()
```


# Build steps (observed and available) and the 4 models in 2011
```{r eval=FALSE, include=FALSE}
# initialisation
steps <- list()
models <- list()
# env_issa <- list()
```

## Running iSSA for shearwaters (including some covariates NA)
```{r eval=FALSE, include=FALSE}
# running the analysis for 2011 data
start <- Sys.time()
for (site in unique(ALL$Site)){
  print(site)
  stps <- iSSA_steps(colony = site, n_control = 6,
                     covariates = c("Bathy", "GBathy", "SST28", "GSST28","Var_SST28",
                                    "SLA1", "VEL7", "GCHLA28"))

  # # removing observed + available steps if NA
  # na.ind <- unique(which(is.na(stps), arr.ind = TRUE)[,1])
  # na.step_id <- stps[na.ind, "step_id_"] %>% unique() %>% pull()
  # stps <- stps %>% 
  #   filter(!step_id_%in%na.step_id)
  
  # saving steps dataframe
  steps[[site]] <- stps

  # saving models
  models[[site]] <- stps %>% amt::fit_issf(case_ ~ sl_ + ta_ + log_sl + Bathy +
                                  SST28 + GSST28 + Var_SST28 + SLA1 + VEL7 +
                                  GCHLA28 + strata(step_id_))
}
end <- Sys.time()
print(end-start)

save(models, steps, file = "issa_env_n6.RData")
```

## Running iSSA for 2011 shearwaters (witout oceanographics NA steps)
```{r eval=FALSE, include=FALSE}
# # running the analysis for 2011 data
# start <- Sys.time()
# for (site in unique(ALL$Site)){
#   print(site)
#   stps <- iSSA_steps(colony = site, year = 2011,
#                       covariates = c("Bathy", "GBathy", "SST28", "GSST28", "Var_SST28",
#                                      "SLA1", "VEL7", "logCHLA28", "GCHLA28"))
# 
#   # removing observed + available steps if NA
#   na.ind <- unique(which(is.na(stps), arr.ind = TRUE)[,1])
#   na.step_id <- stps[na.ind, "step_id_"] %>% unique() %>% pull()
#   stps <- stps %>%
#     filter(!step_id_%in%na.step_id)
#   
#   # saving steps dataframe
#   steps[[site]] <- stps
# 
#   # saving models
#   models[[site]] <- stps %>% amt::fit_issf(case_ ~ sl_ + ta_ + log_sl + Bathy + GBathy +
#                                   SST28 + GSST28 + Var_SST28 + SLA1 + VEL7 +
#                                   logCHLA28 + GCHLA28 + strata(step_id_))
# }
# end <- Sys.time()
# print(end-start)
# 
# env_issa[["clean"]] <- list(steps, models)
```

<!-- ## Running iSSA for 2012 shearwaters -->
```{r eval=FALSE, include=FALSE}
# # running the analysis for 2012 data
# start <- Sys.time()
# for (site in unique(ALL2012$Site)){
#   print(site)
#   stps <- iSSA_steps(colony = site, year = 2012,
#                       covariates = c("Bathy", "GBathy", "SST28", "GSST28", "Var_SST28",
#                                      "SLA1", "VEL7", "logCHLA28", "GCHLA28"))
#   
#   # removing observed + available steps if NA
#   na.ind <- unique(which(is.na(stps), arr.ind = TRUE)[,1])
#   na.step_id <- stps[na.ind, "step_id_"] %>% unique() %>% pull()
#   stps <- stps %>% 
#     filter(!step_id_%in%na.step_id)
#   
#   # saving steps dataframe
#   steps[[site]] <- stps
#   
#   # saving models
#   models[[site]] <- stps %>% amt::fit_issf(case_ ~ sl_ + ta_ + log_sl + Bathy + GBathy +
#                                   SST28 + GSST28 + Var_SST28 + SLA1 + VEL7 +
#                                   logCHLA28 + GCHLA28 + strata(step_id_))
# }
# end <- Sys.time()
# print(end-start)
# 
# env_year[[2012]] <- list(steps, models)
```


## Load models and steps generated before
```{r}
# load("issa_env.RData")
load("issa_env_n6.RData")
```


<!-- compare Lavezzi birds'positions to remaining observed steps -->
<!-- ```{r} -->
<!-- obs_stps_miss_L <- steps_miss$Lavezzi %>% filter(case_ == TRUE) -->

<!-- mediterranean  <- make_bbox(lat = y2_, lon = x2_, data = obs_stps_miss_L) -->
<!-- map_steps <- get_map(location = mediterranean) %>% ggmap() -->
<!-- map_steps + geom_point(data = obs_stps_miss_L, aes(x = x2_, y = y2_), size = 0.1) -->

<!-- mediterranean  <- make_bbox(lat = Latitude, lon = Longitude, -->
<!--                             data = ALL2012[ALL2012$Site=="Lavezzi",]) -->
<!-- map_coord <- get_map(location = mediterranean) %>% ggmap() -->
<!-- map_coord + geom_point(data = ALL2012[ALL2012$Site=="Lavezzi",], -->
<!--                        aes(x = Longitude, y = Latitude), size = 0.1) -->
<!-- ``` -->


Correlation analysis for the covariates
```{r}
grd <- grid_oceano_2011[[36]]
covariates_corr <- cor(na.omit(grd[, c("Bathy", "GBathy", "SST28", "GSST28", "Var_SST28",
                                    "SLA1", "VEL7", "logCHLA28", "GCHLA28")]))
res.PCA <- FactoMineR::PCA(na.omit(grd[, c("Bathy", "GBathy", "SST28", "GSST28", "Var_SST28",
                                    "SLA1", "VEL7", "logCHLA28", "GCHLA28")]))
corrplot::corrplot(covariates_corr)
```



```{r}
colony <- "Riou"
print(colony)
modR <- models$Riou$model ; modR
```

```{r echo=FALSE, results='asis'}
stargazer(modR, type = "latex",
          title = paste("Résumé du modèle de régression logistique conditionnel pour la colonie", colony))
```

```{r message=FALSE, warning=FALSE}
obs_stps <- steps$Riou %>% filter(case_ == TRUE)
rdm_stps <- steps$Riou %>% filter(case_ == FALSE)

mediterranean  <- make_bbox(lat = y2_, lon = x2_, data = steps$Riou)
map <- get_map(location = mediterranean) %>% ggmap()

png(paste0("figures/", colony, "_map.png"), width=4.4, height=4, units="in", res=200)
map + geom_point(data = rdm_stps, aes(x = x2_, y = y2_),
             size = 0.1, color = "red", alpha = 0.1) +
  geom_point(data = obs_stps, aes(x = x2_, y = y2_), size = 0.1) +
  ggtitle(colony)
dev.off()
```


```{r}
colony <- "Lavezzi"
print(colony)
modL <- models$Lavezzi$model ; modL
```

```{r echo=FALSE, results='asis'}
stargazer(modL, type = "latex", title = paste(
  "Résumé du modèle de régression logistique conditionnel pour la colonie",
  colony), summary = FALSE)
```

There are too many steps removed because missing. Let's investigate why.
```{r message=FALSE, warning=FALSE}
obs_stps <- steps$Lavezzi %>% filter(case_ == TRUE)
rdm_stps <- steps$Lavezzi %>% filter(case_ == FALSE)

mediterranean  <- make_bbox(lat = y2_, lon = x2_, data = steps$Lavezzi)
map <- get_map(location = mediterranean) %>% ggmap()

png(paste0("figures/", colony, "_map.png"), width=4.4, height=4, units="in", res=200)
map + geom_point(data = rdm_stps, aes(x = x2_, y = y2_),
             size = 0.1, color = "red", alpha = 0.1) +
  geom_point(data = obs_stps, aes(x = x2_, y = y2_), size = 0.1) +
  ggtitle(colony)
dev.off()
```

Taille des vols
```{r}
steps$Lavezzi %>% as_tibble() %>% 
  group_by(burst_) %>% 
  summarise(nb_stps = n_distinct(step_id_)) %>% 
  filter(nb_stps < 10)
```



```{r}
colony <- "Porquerolles"
print(colony)
modP <- models$Porquerolles$model ; modP
```

```{r echo=FALSE, results='asis'}
stargazer(modP, type = "latex", title = paste(
  "Résumé du modèle de régression logistique conditionnel pour la colonie",
  colony), summary = FALSE)
```

```{r message=FALSE, warning=FALSE}
obs_stps <- steps$Porquerolles %>% filter(case_ == TRUE)
rdm_stps <- steps$Porquerolles %>% filter(case_ == FALSE)

mediterranean  <- make_bbox(lat = y2_, lon = x2_, data = steps$Porquerolles)
map <- get_map(location = mediterranean) %>% ggmap()

png(paste0("figures/", colony, "_map.png"), width=4.4, height=4, units="in", res=200)
map + geom_point(data = rdm_stps, aes(x = x2_, y = y2_),
             size = 0.1, color = "red", alpha = 0.1) +
  geom_point(data = obs_stps, aes(x = x2_, y = y2_), size = 0.1) +
  ggtitle(colony)
dev.off()
```


```{r}
colony <- "Giraglia"
print(colony)
modG <- models$Giraglia$model ; modG
```

```{r echo=FALSE, results='asis'}
stargazer(modG, type = "latex", title = paste(
  "Résumé du modèle de régression logistique conditionnel pour la colonie",
  colony), summary = FALSE)
```

```{r message=FALSE, warning=FALSE}
obs_stps <- steps$Giraglia %>% filter(case_ == TRUE)
rdm_stps <- steps$Giraglia %>% filter(case_ == FALSE)

mediterranean  <- make_bbox(lat = y2_, lon = x2_, data = steps$Giraglia)
map <- get_map(location = mediterranean) %>% ggmap()

png(paste0("figures/", colony, "_map.png"), width=4.4, height=4, units="in", res=200)
map + geom_point(data = rdm_stps, aes(x = x2_, y = y2_),
             size = 0.1, color = "red", alpha = 0.1) +
  geom_point(data = obs_stps, aes(x = x2_, y = y2_), size = 0.1) +
  ggtitle(colony)
dev.off()
```


Taille des vols
```{r}
steps$Lavezzi %>% as_tibble() %>% 
  group_by(burst_) %>% 
  summarise(nb_stps = n_distinct(step_id_)) %>% 
  filter(nb_stps < 10)
```




# Utilization distribution

## Create a grid from the centroid coordinates

```{r eval=FALSE, include=FALSE}
# # Create an sf object from a data frame of more than two columns
# x <- grd[,c("Longitude", "Latitude","Bathy", "GBathy", "SST28", "GSST28", "Var_SST28",
#                                      "SLA1", "VEL7", "logCHLA28", "GCHLA28")]
# poly <- eSDM::pts2poly_centroids(x, 0.0675/2, crs = 2154, agr = "constant")
# 
# mediterranean  <- make_bbox(lat = Latitude, lon = Longitude, data = grd)
# map <- get_map(location = mediterranean) %>% ggmap()
# 
# # medit <- c(left = min(grd$Longitude),
# #            bottom = min(grd$Latitude),
# #            right = max(grd$Longitude),
# #            top = max(grd$Latitude))
# # map <- get_stamenmap(medit, zoom = 7, "terrain-background") %>% ggmap()
# 
# 
# world <- ne_countries(scale = "medium", returnclass = "sf")
# ggplot(poly) + 
#   geom_sf(aes(fill = Bathy), color = NA)
```

